\documentclass[a4paper,12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb,amsthm,graphicx}
%\usepackage{bbm} (Do not want to use these, may create conflicts)
%\usepackage{gensymb}
\usepackage{enumitem}
\usepackage{color}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{pdfpages}
\usepackage{subcaption}
\usepackage[font=small]{caption}
\usepackage[hang,flushmargin]{footmisc} 
\usepackage{placeins}
\usepackage{float}
\usepackage{rotating,tabularx}
\usepackage{booktabs}
\usepackage[mathscr]{euscript}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{mathrsfs}
\usepackage[left=2.7cm,right=2.7cm,bottom=2.7cm,top=2.7cm]{geometry}
\parindent0pt 

\input{macros}



\begin{document}



\heading{Multiscale Inference}{and Long-Run Variance Estimation}{in Nonparametric Regression}{with Time Series Errors}

%\vspace{-0.5cm}

\authors{Marina Khismatullina\renewcommand{\thefootnote}{1}\footnotemark[1]}{University of Bonn}{Michael Vogt\renewcommand{\thefootnote}{2}\footnotemark[2]}{University of Bonn} 
\footnotetext[1]{Address: Bonn Graduate School of Economics, University of Bonn, 53113 Bonn, Germany. Email: \texttt{marina.k@uni-bonn.de}.}
\renewcommand{\thefootnote}{2}
\footnotetext[2]{Corresponding author. Address: Department of Economics and Hausdorff Center for Mathematics, University of Bonn, 53113 Bonn, Germany. Email: \texttt{michael.vogt@uni-bonn.de}.}
\renewcommand{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{2}

\vspace{-0.5cm}

\version{\today}

\vspace{-1.2cm}



\renewcommand{\abstractname}{}
\begin{abstract}
\noindent In this paper, we develop new multiscale methods to test qualitative hypotheses about the function $m$ in the nonparametric regression model $Y_{t,T} = m(t/T) + \varepsilon_t$ with time series errors $\varepsilon_t$. In time series applications, $m$ represents a nonparametric time trend. Practitioners are often interested in whether the trend $m$ has certain shape properties. For example, they would like to know whether $m$ is constant or whether it is increasing/decreasing in certain time regions. Our multiscale methods allow to test for such shape properties of the trend $m$. In order to perform the methods, we require an estimator of the long-run error variance $\sigma^2 = \sum\nolimits_{\ell=-\infty}^{\infty} \cov(\varepsilon_0,\varepsilon_{\ell})$. We propose a new difference-based estimator of $\sigma^2$ for the case that $\{ \varepsilon_t \}$ is an AR($p$) process. In the technical part of the paper, we derive asymptotic theory for the proposed multiscale test and the estimator of the long-run error variance. The theory is complemented by a simulation study and an empirical application to climate data. 
\end{abstract}


\renewcommand{\baselinestretch}{1.2}\normalsize

\textbf{Key words:} Multiscale statistics; long-run variance; nonparametric regression; time series errors; shape constraints; strong approximations; anti-concentration bounds.

\textbf{AMS 2010 subject classifications:} 62E20; 62G10; 62G20; 62M10. 

\numberwithin{equation}{section}
\allowdisplaybreaks[1]



\input{paper_intro}
\input{paper_methods}
\input{paper_app}

\bibliographystyle{ims}
{\small
\setlength{\bibsep}{0.55em}
\bibliography{bibliography}}



\end{document}
